---
title: "Customizing a Website"
---

:::{.callout-tip icon="false" collapse="false"}
### {{< fa brain >}} Learning Objectives

By the end of this module, you will be able to:

- <u>Identify</u> components of SCSS stylesheets
- <u>Describe</u> key code syntax differences between SCSS and R code
- <u>Use</u> SCSS stylesheets to customize a deployed Quarto website

:::

## SCSS Background

Technically, when we render a Quarto project we are asking Quarto to convert the `.qmd` file into `.html` and apply a CSS (<u>C</u>ascading <u>S</u>tyle <u>S</u>heets) stylesheet. This is why your deployed website up to this point has had a blue navbar, a white background, and black text (to name a few theme elements governed by the implied CSS stylesheet).

Quarto supports us making our own, customized stylesheets with [Sass](https://sass-lang.com/) (<u>S</u>yntactically <u>A</u>wesome <u>S</u>tyle<u>s</u>heet) to tailor the look and feel of our website to exactly our personal preference. The relevant file type is `.scss` (<u>S</u>assy CSS; still pronounced "sass"). These files can be written in a combination of CSS and Sass so the syntax might look a little alien to you but this workshop will try to provide a useful level of practical detail for you to engage with these stylesheets even if this is your first foray into that world.

:::{.callout-note}
#### R Versus SCSS Comparison

If you're an R coder, it might be helpful to compare how R assigns values to objects versus how SCSS binds values to variables as well as how to write a comment (i.e., a non-coding line).

{{< fa brands r-project >}} R code
```
# Comment
object <- value
```

{{< fa brands sass >}} Sass code
```
// Comment
$object: value;
```
:::

## Customize the Website with Sass

### 1. Create Sass Stylesheets

**Create two new files: "theme_light.scss" and "theme_dark.scss".** This will let us easily create a light mode and a dark mode for our website so that visitors can click a toggle in the navbar to choose their desired aesthetic.

Once you have these files, **click the collapsed menu below and copy/paste all of its contents into both empty SCSS files.**

<details> 
<summary>SCSS Contents</summary>

```{r}
#| label: customize-scss
#| eval: false

/*-- scss:defaults --*/

// Colors
/* Found on https://coolors.co/palettes/trending */
$org-1: #ff9f1c;
$org-2: #ffbf69;
$mint-1: #2ec4b6;
$mint-2: #cbf3f0;
$url-blue: #4361ee;

/* generally useful other colors */
$cream: #F4F3EE;
$charcoal: #191919;
$gray-1: #8A817C;
$gray-2: #BCB8B1;
$gray-3: #d6d6d6;

// Fonts
$font-size: 25px;
@import url('https://fonts.googleapis.com/css2?family=Open+Sans:ital,wght@0,300..800;1,300..800&display=swap');

// Base document colors
$navbar-bg: $org-1; // navbar
$navbar-fg: $charcoal; // navbar foreground elements
$navbar-hl: $org-2; // highlight color when hovering on a page
$sidebar-bg: $org-2; // sidebar
$sidebar-fg: $charcoal; // sidebar foreground elements
$body-bg: $cream; // page background
$body-color: $charcoal; // page text
$footer-bg: $mint-1; // footer
$footer-fg: $charcoal; // footer text
$link-color: $url-blue; // hyperlinks

// Code text
$code-bg: $gray-2; // inline code background color
$code-color: $charcoal; // inline code text color
$code-block-bg: $gray-2; // code block background color

/*-- scss:rules --*/

// Underline URLs
.reveal .slide a { text-decoration: underline; }

// Improve background contrast for code chunk ouput
pre code:not(.sourceCode) { background-color: $gray-2; }
```

</details>

This will result in your light and dark mode being identical to start with but that's okay because this method makes sure that the format of both SCSS files is correct.

### 2. Integrate Into the Website

To get your website to use these `.scss` files, we'll need to make some minor edits to the website's YAML. **Open your "_quarto.yml" file in your IDE and add the following lines**.

:::{.callout-warning}
#### Paste in the Right Place!

Note that the below code chunk includes a `html:` element but your existing `_quarto.yml` _will already have that element!_ That bit is included below so you know _where_ to copy/paste the new lines into your existing YAML.

If you're using a mostly unmodified `_quarto.yml`, this will be about 4 lines up from the very bottom.
:::

```{r}
#| label: global-yaml-scss
#| eval: false
  html:
    theme:
      light: theme_light.scss
      dark: theme_dark.scss
    mainfont: Open Sans
```

**Double check that your indentation is right after you copy/paste those lines!** `theme:` and `mainfont` should both be one indent more than `html:` while `light:` and `dark:` should have an additional indent beyond `theme:`.

### 3. Customize to Your Taste

The Sass code you copy/pasted is functional and inoffensive but likely does not reflect your personal aesthetic sensibilities. So, take some time to find some colors that you prefer. This is also a great time to make your dark mode actually a dark mode (currently it has the same content as the light mode `.scss` file).

:::{.panel-tabset}
#### Customizing Colors

First, I'd recommend visiting [coolors.co](https://coolors.co/palettes/popular) and looking at their trending color palettes for one that feels like your vibe. To incorporate that in your site **you'll need to copy the hexadecimal code (6-digit color code) and add it to the relevant `.scss` file**.

Typically the simplest way to do this is to **assign each color to a variable name**, just like the default colors you copy/pasted in earlier (e.g., `$cream: #F4F3EE;`). Be sure that each line starts with a `$` and ends with a `;`.

Once you've assigned variable names to your new color(s), **update the existing fields in the lower half of the `.scss` file** with those variable names. Consult the comments for some explanation of what those variable names are called. Remember too that each variable name needs to start with `$` when you invoke it (so that Sass knows you're talking about an existing variable). 

#### Customizing Fonts

To find new possible fonts, check out Google Fonts ([fonts.google.com](https://fonts.google.com/)) and search around for one that seems like a good fit for your personal style. When you have identified a font that you like, follow these instructions:

1. Click on the font to go to its homepage
2. Click the **<span style="color:blue"/>blue</span>** "Get font" button in the top right corner of the page
3. Click the **<span style="color:blue"/>blue</span>** "{{< fa code >}} Get embed code" button
4. Click the `@import` radiobutton in the top of the resulting sidebar
5. In the first code chunk, copy everything between the two `<style>` HTML tags
6. In your `.scss` files, paste the copied text somewhere near the font that is already there to begin with
7. Finally, update the `mainfont:` element of `_quarto.yml` with the new font name

:::

### 4. Test it Locally

To make sure that your new website displays correctly, run {{< fa terminal >}} `quarto preview`. If it doesn't, edit the offending file(s), save those edits, and look at the preview again.

This step can be especially critical when fiddling with (yes, that's the technical term {{< fa smile >}}) colors for your website. Use the preview extensively to be sure you're happy with your chosen color palette.

### 5. Render & Push

Once you're happy with how the preview looks:

1. Run {{< fa terminal >}} `quarto render`
2. Commit & push/sync all the files that you changed
3. Wait for the GHA to complete

## Activity - Try it Out

:::{.callout-note}
### Your Turn!

Let's take a bit for you to customize your Sass stylesheets!
:::

## Additional Information

- UCSB Master of Environmental Data Science (MEDS) "[Customizing Quarto Websites](https://ucsb-meds.github.io/EDS-296-DS-portfolios/course-materials/week2-customizing-websites.html)" lesson
    - [Lecture slides](https://ucsb-meds.github.io/customizing-quarto-websites/#/title-slide) for that lesson
